---
title: "JOTW data analysis"
output: html_document
---

```{r setup, include=FALSE}
library('genderizeR')
library('tidyr')
library('data.table')
library('ggplot2')
library('scales')
library('RColorBrewer')
library('googlesheets')
```

# Activity Over Time

First, we analyzed the number of reactions and posts over time.

```{r}
statusesSheet <- gs_title("Statuses")
statuses <- statusesSheet %>% gs_read(ws = 'Statuses') 
statuses <- statuses[!is.na(statuses$status_published),]
statuses$yrmonth = paste(year(as.Date(statuses$status_published, format = "%m/%d/%Y")), 
    strftime(as.Date(statuses$status_published, format = "%m/%d/%Y"), "%m"), sep = "-")
ggplot(aes(x = statuses$yrmonth, y = statuses$num_reactions), data = statuses) + 
  geom_bar(stat= "identity") + 
  labs(title = "Reaction activity level over time", 
      x = "Time (Year-Month)", 
      y = "Number of Reactions")
ggplot(aes(x = statuses$yrmonth), data = statuses) + 
  geom_bar() + 
  labs(title = "Post activity level over time", 
    x = "Time (Year-Month)", 
    y = "Number of Posts") 
```
# Activity Over Percentiles

Next, we analyzed the average number of statuses made by the top .1, 1, 5 and 10 percent of users.

```{r pressure, echo=FALSE}
#aggpostdata is a data frame with two columns (full name and total number of posts)
aggpostdata <-aggregate(statuses, by=list(statuses$status_author), FUN = length)
aggpostdata <- aggpostdata[order(aggpostdata$status_id),]
aggpostdata$name <- aggpostdata$Group.1
aggpostdata$postcount = aggpostdata$status_id
aggpostdata <- aggpostdata[c(20:21)]
```


```{r pressure, echo=FALSE}
name = c("Top0.1", "Top01", "Top05", "Top10", "Top20", "everyone")
mean = c(mean(aggpostdata[aggpostdata$postcount > quantile(aggpostdata$postcount,prob=1-.1/100),]$postcount), 
    mean(aggpostdata[aggpostdata$postcount > quantile(aggpostdata$postcount,prob=1-1/100),]$postcount), 
    mean(aggpostdata[aggpostdata$postcount > quantile(aggpostdata$postcount,prob=1-5/100),]$postcount), 
    mean(aggpostdata[aggpostdata$postcount > quantile(aggpostdata$postcount,prob=1-10/100),]$postcount), 
    mean(aggpostdata[aggpostdata$postcount > quantile(aggpostdata$postcount,prob=1-20/100),]$postcount), 
    mean(aggpostdata$postcount))
pop = c(nrow(aggpostdata[aggpostdata$postcount > quantile(aggpostdata$postcount,prob=1-.1/100),]), 
    nrow(aggpostdata[aggpostdata$postcount > quantile(aggpostdata$postcount,prob=1-1/100),]), 
    nrow(aggpostdata[aggpostdata$postcount > quantile(aggpostdata$postcount,prob=1-5/100),]), 
    nrow(aggpostdata[aggpostdata$postcount > quantile(aggpostdata$postcount,prob=1-10/100),]), 
    nrow(aggpostdata[aggpostdata$postcount > quantile(aggpostdata$postcount,prob=1-20/100),]), 
    nrow(aggpostdata))
df = data.frame(name, mean, pop)
ggplot(df, aes(name, mean)) + 
  geom_bar(stat = "identity") + 
  labs(title = "Average Number of Posts by Percentile", 
    x = "Group of Users", 
    y = "Average Number of Posts") + 
  geom_text(aes(label=pop), vjust= -.2)
```
# General Activity Analysis

Then, we analyzed user activity through a 2D histogram. We wanted to know the distribution of people who post and those 
who comment. We found that a lot of people in the JOTW facebook group write one or two comments and don't post.

```{r pressure, echo=FALSE}
commentspt1Sheet <- gs_title("Comments1")
commentspt1 <- commentspt1Sheet %>% gs_read(ws = 'Comments')
commentspt2Sheet <- gs_title("Comments2")
commentspt2 <- commentspt2Sheet %>% gs_read(ws = 'Comments')

comments <- rbind(commentspt1, commentspt2)
rm(commentspt1)
rm(commentspt2)

#aggcommentdata is a data frame with two columns (full name and total number of comments)
aggcommentdata <-aggregate(comments, by=list(comments$comment_author), FUN = length)
aggcommentdata <- aggcommentdata[order(aggcommentdata$comment_id),]
aggcommentdata$name <- aggcommentdata$Group.1
aggcommentdata$commentcount = aggcommentdata$status_id
aggcommentdata <- aggcommentdata[order(aggcommentdata$name),]
aggcommentdata <- aggcommentdata[c(15:16)][-c(1:8), ]
```

```{r pressure, echo=FALSE}
#aggdata is a data frame with three columns (full name, total number of comments and total number of posts)
aggdata <- merge(x = aggcommentdata, y = aggpostdata, by = "name", all = TRUE)
aggdata[is.na(aggdata)] <- 0.1 #NA values are changed to be 0.1
aggdata <- aggdata[order(aggdata$postcount),] 
rf <- colorRampPalette(rev(brewer.pal(11,'Spectral')))
r <- rf(32)
ggplot(aggdata, aes(aggdata$postcount,aggdata$comment)) +
  stat_bin2d(bins=25) + 
  scale_x_log10() + 
  scale_y_log10() + 
  scale_fill_gradientn(colors=r, trans="log") + 
  labs(title = "2D Histogram of User Activity", 
    x = "Number of Posts (Logarithmically Scaled)", 
    y = "Number of Comments (Logarithmically Scaled)")
aggdata[aggdata==0.1] <- 0 #revert 0.1 values back to 0 values
```

```{r pressure, echo=FALSE}
aggdata$activity <- ifelse(aggdata$commentcount>0&aggdata$postcount>0, "Both", ifelse(aggdata$commentcount>0, 
    "Comment Only", "Post Only"))
```

#QQ Plots

Then, we created QQ plots to see what percent of users account for what percent of posts, reacts and comments.

```{r pressure, echo=FALSE}
#add "num_reactions",	"num_comments",	"num_shares", "num_likes", "num_loves",	"num_wows",	"num_hahas", "num_sads", 
#"num_angrys" columns to aggdata data frame
aggdata$Group.1 = aggdata$name
aggdata <- merge(x = aggregate(statuses[,c("num_reactions",	"num_comments",	"num_shares", "num_likes", "num_loves",
	"num_wows",	"num_hahas", "num_sads", "num_angrys")], by=list(statuses$status_author), "sum", na.rm=TRUE),
	y = aggdata, by = "Group.1", all = TRUE)
```


```{r pressure, echo=FALSE}
aggdata <- aggdata[order(aggdata$postcount),] 
x = subset(aggdata, postcount > 0)$postcount
n = length(x)
qqplot((1:n - 1)/(n - 1), cumsum(subset(aggdata, postcount > 0)$postcount)/
    sum(subset(aggdata, postcount > 0)$postcount),  xlab = "% of Users", ylab = "% of Posts", 
    main = "QQ Plot: % of Users vs % of Posts Made")
```
```{r pressure, echo=FALSE}
aggdata <- aggdata[order(aggdata$commentcount),] 
x = subset(aggdata, commentcount > 0)$commentcount
n = length(x)
qqplot((1:n - 1)/(n - 1), cumsum(subset(aggdata, commentcount > 0)$commentcount)/
    sum(subset(aggdata, commentcount > 0)$commentcount), xlab = "% of Users", ylab = "% of Comments", 
    main = "QQ Plot: % of Users vs % of Comments Made")
```

```{r pressure, echo=FALSE}
aggdata <- aggdata[order(aggdata$num_reactions),] 
x = aggdata[!is.na(aggdata$num_reactions),]$num_reactions
n = length(x)
qqplot((1:n - 1)/(n - 1),cumsum(aggdata[!is.na(aggdata$num_reactions),]$num_reactions)
    /sum(aggdata[!is.na(aggdata$num_reactions),]$num_reactions), xlab = "% of Users", ylab = "% of Reacts", 
    main = "QQ Plot: % of Users vs % of Reacts Received")
```

# Activity By Gender

Afterward, we conducted basic distribution analysis for gender.

```{r pressure, echo=FALSE}
#add "comm_num_reactions", "comm_num_likes", "comm_num_loves", "comm_num_wows",	"comm_num_hahas", "comm_num_sads", 
#"comm_num_angrys" columns to aggdata data frame
aggdata = merge(x = setNames(aggregate(comments[,c("num_reactions",	"num_likes", "num_loves", "num_wows", "num_hahas", 
    "num_sads", "num_angrys")], by=list(comments$comment_author), "sum", na.rm=TRUE), c("Group.1", "comm_num_reactions", 
    "comm_num_likes", "comm_num_loves",	"comm_num_wows",	"comm_num_hahas", "comm_num_sads", "comm_num_angrys")), 
    y = aggdata, by = "Group.1", all = TRUE)
aggdata = aggdata[c(2:21)]
```

```{r pressure, echo=FALSE}
#Set fullname column to full name and name column to first name
aggdata$fullname = aggdata$name
aggdata$name = sub(" .*", "", aggdata$name)
```

```{r pressure, echo=FALSE}
#Merged is a data frame with four columns (first name, probability, count and gender)

#givenNames = findGivenNames(unique(sub(" .*", "", aggdata$name)), apikey = "8851672c0e70afb02af3bc2c479a7236")
#genderedData <- as.data.frame(genderize(unique(aggdata$name), genderDB=givenNames, blacklist=NULL))
#genderedData$name <- genderedData$givenName
#merged <- merge(x = givenNames, y = genderedData, by = "name", all = TRUE)

#rm(genderedData)
#rm(givenNames)

#merged$gender <- merged$gender.y
#merged$name <- merged$text
#merged <- merged[, c("name", "probability", "count", "gender")]
#merged <- subset(merged[!is.na(merged$name),])

#We don't want to run genderize.io again so use the csv!!!
merged = 
as.data.frame(fread('GenderData.csv',header=TRUE,encoding='UTF-8',fill=TRUE))[c(-1)]
```

```{r pressure, echo=FALSE}
#Threshold tests
merged$M <- as.numeric(merged$probability, na.rm = TRUE) * as.numeric(merged$count, na.rm = TRUE) + 3.5
merged$F <- (1 - as.numeric(merged$probability, na.rm = TRUE)) * as.numeric(merged$count, na.rm = TRUE) + 3.5
merged$percent <- as.numeric(merged$F, na.rm = TRUE) / (as.numeric(merged$M, na.rm = TRUE) 
    + as.numeric(merged$F, na.rm = TRUE))

#Drops data points that don't have a percentage > 95% or < 5%
merged <- subset(merged[!is.na(merged$gender),], percent >= .95 | percent <= 0.05)

```

```{r pressure, echo=FALSE}
aggdata$gender <- NA
aggdata$probability <- NA
aggdata$count <- NA
aggdata <- merge(x = aggdata, y = merged, by = "name", all = TRUE)
aggdata$gender <- aggdata$gender.y
aggdata$probability <- aggdata$probability.y
aggdata$count <- aggdata$count.y
aggdata <- aggdata[!is.na(aggdata$name),][c(-1, -22:-30)]
```

```{r pressure, echo=FALSE}
barplot(table(aggdata$gender), main="Barplot of Gender", xlab="gender")
```

```{r pressure, echo=FALSE}
#Overall activity
barplot(table(aggdata$activity), main="Barplot of User Activity", xlab = "activity", ylab = "Frequency")
ggplot(aggdata, aes(x = activity, ..count.., group=gender)) + 
                geom_bar(aes(fill = gender, 
                             y = c(..count..[..group..==1]/sum(..count..[..group..==1]),           
                                   ..count..[..group..==2]/sum(..count..[..group..==2]), 
                                   ..count..[..group..==3]/sum(..count..[..group..==3]))), 
                         position = "dodge") + 
  scale_y_continuous(labels = percent) + 
  labs(title = "Histogram of User Activity by Gender", x = "Activity Level", y = "Percentage")
```

```{r pressure, echo=FALSE}
#change to something more appropriate
aggdata$postbin <- cut(aggdata$postcount, breaks = c(0, 1, 5, 100, 1134), include.lowest = TRUE, right = FALSE)
aggdata$commentbin <- cut(aggdata$commentcount, breaks = c(0, 1, 5, 10, 50, 5168), include.lowest = TRUE, right = FALSE) 
aggdata$reactbin <- cut(aggdata$num_reactions, breaks = c(0, 1, 5, 10, 50, 54224))
```

```{r pressure, echo=FALSE}
#Count histogram
ggplot(aggdata, aes(x = postbin, group=gender)) + 
       geom_bar(aes(fill = gender), position = "dodge") +
       theme_bw() +
       labs(title = "Histogram of User Activity by Gender", 
       x = "Number of Posts Made", 
       y = "Count")
```

```{r pressure, echo=FALSE}
#Binned percentage histogram
ggplot(aggdata, aes(x = postbin, ..count.., group=gender)) + 
       geom_bar(aes(fill = gender, 
       y = c(..count..[..group..==1]/sum(..count..[..group..==1]),           
                                   ..count..[..group..==2]/sum(..count..[..group..==2]), 
                                   ..count..[..group..==3]/sum(..count..[..group..==3]))), 
       position = "dodge") + 
       scale_y_continuous(labels = percent) + 
       theme_bw() +
       labs(title = "Histogram of User Activity by Gender", 
       x = "Number of Posts Made", 
       y = "Percentage")
```

```{r pressure, echo=FALSE}
#Count histogram
ggplot(aggdata, aes(x = commentbin, ..count.., group=gender)) + 
       geom_bar(aes(fill = gender), position = "dodge") +
       theme_bw() +
       labs(title = "Histogram of User Activity by Gender", 
       x = "Number of Comments Made", 
       y = "Count")
```

```{r pressure, echo=FALSE}
#Binned percentage histogram
ggplot(aggdata, aes(x = commentbin, ..count.., group=gender)) + 
                geom_bar(aes(fill = gender, 
                             y = c(..count..[..group..==1]/sum(..count..[..group..==1]),           
                                   ..count..[..group..==2]/sum(..count..[..group..==2]), 
                                   ..count..[..group..==3]/sum(..count..[..group..==3]))), 
                         position = "dodge") + 
        scale_y_continuous(labels = percent)  + 
        theme_bw() +
        labs(title = "Histogram of User Activity by Gender", 
        x = "Number of Comments Made", 
        y = "Percentage")
```
```{r pressure, echo=FALSE}
#Count histogram
ggplot(aggdata[!is.na(aggdata$reactbin),], aes(x = reactbin, ..count.., group=gender)) + 
        geom_bar(aes(fill = gender), position = "dodge") +
        theme_bw() +
        labs(title = "Histogram of User Feedback by Gender", 
        x = "Number of Reacts (TOTAL)", 
        y = "Count")
```

```{r pressure, echo=FALSE}
#Binned percentage histogram
ggplot(aggdata[!is.na(aggdata$reactbin),], aes(x = reactbin, ..count.., group=gender)) + 
                geom_bar(aes(fill = gender, 
                             y = c(..count..[..group..==1]/sum(..count..[..group..==1]),           
                                   ..count..[..group..==2]/sum(..count..[..group..==2]), 
                                   ..count..[..group..==3]/sum(..count..[..group..==3]))), 
                         position = "dodge") + 
        scale_y_continuous(labels = percent) + 
        theme_bw() + 
        labs(title = "Histogram of User Feedback by Gender", 
        x = "Number of Reacts (TOTAL)", 
        y = "Percentage")
```

```{r pressure, echo=FALSE}
aggdata$reactbin <- cut(aggdata$num_reactions / aggdata$postcount, breaks = c(0, 1, 5, 10, 30, 3800))
#Count histogram
ggplot(aggdata[!is.na(aggdata$reactbin),], aes(x = reactbin, ..count.., group=gender)) + 
        geom_bar(aes(fill = gender), position = "dodge") +
        theme_bw() +
        labs(title = "Histogram of User Feedback by Gender", 
        x = "Number of Reacts (PER POST)", 
        y = "Count")
```

```{r pressure, echo=FALSE}
#Binned percentage histogram
ggplot(aggdata[!is.na(aggdata$reactbin),], aes(x = reactbin, ..count.., group=gender)) + 
                geom_bar(aes(fill = gender, 
                             y = c(..count..[..group..==1]/sum(..count..[..group..==1]),           
                                   ..count..[..group..==2]/sum(..count..[..group..==2]), 
                                   ..count..[..group..==3]/sum(..count..[..group..==3]))), 
                         position = "dodge") + 
        scale_y_continuous(labels = percent) + 
        theme_bw() + 
        labs(title = "Histogram of User Feedback by Gender", 
        x = "Number of Reacts (PER POST)", 
        y = "Percentage")
```